---
title: "Practice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages ('lmtest')
#install.packages ('broom.mixed') - so our glance function works.
library (lmtest)
library (tidyverse)
library (broom)
library (nlme)
library (broom.mixed)
```

Don´t have hashtags install.package on your markdown because it will install several times.

#Spatial and temporal structure
Richard J. Telford
June 1, 2018

1. Simulate a 100-observation autocorrelated timeseries with arima.sim, with a first order autoregressive coefficient of 0.5. Also make a time vector of 1:100


2. Plot the data.
We don´t use c because we just have one value. 
```{r}
Sim <- arima.sim (n=100, list(ar = 0.5))
Sim
plot(Sim)
```


3. Regress the timeseries against time with an OLS model. Does the model appear to be statistically significant?

```{r}

set.seed (42)
dat <- tibble(
  x = 1:100,
  y = arima.sim (n= 100, list(ar = 0.5)))

  mod <- lm (y ~ x, data = dat)
  
  anova (mod)
  summary (mod)
  
  par(mfrow = c(2,2))
  plot (mod)
    
    
```


4. Plot the model diagnostics, including an acf (autocorrelation function) and pacf (partial correlation functionof the residuals.

```{r}
acf(resid(mod))
pacf (resid(mod))
```


5. Use the Durbin-Watson test to test the residuals for autocorrelation.

DW = is less than two 

```{r}

dwtest(mod)
```

6. Fit a gls with an appropriate correlation structure. Is this a better model? How have the p-value and effect size changed?
If I type summary for my model in the console, I will have a summarized. Run an anova between a gls with correlation and one gls without the correlation and I then use anova in BOTH to compare which one has a minor AIC, which means that they are more fit. 

```{r}

year <- time (Sim)
fit.gls <- gls(Sim ~ year, correlation = corAR1())
summary (fit.gls)

no.correlation <- gls(Sim~year)
summary (no.correlation)

anova(fit.gls, no.correlation)

```

7. Repeat the above 1000 times and find how autocorrelation affects the distribution of the p-value and the effect size.

Try to color code by significance
Here, I run my model 1000, but my lm has a strong p on the left, as if it were significant; it tests several models and show me what is signifcant. 
With the gls it is corrected, because it shouldn´t be significant, there is only noise, no relation.

in line 108 I used tidy instead of glance, they are not necessarily replaceble but glance didnt work in the gls and I need the numbers it gives. Then I need to filter only for term x, because I don´t want my intercept. 

A gls without any specification work as an lm, so that is why I used it. But gls retrieves some information that lm doesnt, so maybe it is useful to use gls. 

```{r}

rerun(.n=1000, tibble (x = 1:100, y= arima.sim (n=100, list (ar = 0.5)))) %>% 
  map (~ lm(y~x, data = .)) %>% 
  map_df (glance) %>% 
  ggplot (aes(x=p.value)) + geom_histogram() 


rerun(.n=1000, tibble (x = 1:100, y= arima.sim (n=100, list (ar = 0.5)))) %>% 
  map (~ gls(y ~ x, data = ., correlation = corAR1())) %>% 
  map_df (tidy) %>% filter(term == "x") %>% 
  ggplot (aes(x=p.value)) + geom_histogram() 

```


8. Real data
The built-in dataset LakeHuron has annual lake level data from 1885 to 1972 Load the data with the command data(LakeHuron)

```{r}
data (LakeHuron)
```


9. Plot the data.

```{r}
plot (LakeHuron)
```


10. Regress the LakeHuron lake level against year using a linear model. Is there a significant trend?

First, if we retrieve the LakeHuron data we see just a column X with the depth values. But writing in the console, we have the information that this is a time series, year of start and end and we also know the frequency, which is one. From now, before creating a model, we need to create a tibble using tibble() and creating the x with the year(LakeHuron).
Then, I created a lm and ran an ANOVA on it. 
There is a significant trend. p-value: 3.545e-08

```{r}
LakeHuron <- tibble(depth = LakeHuron, year = time(LakeHuron))
LakeHuron
 Lake.mod <- lm ( depth ~ year, data = LakeHuron)
  
  anova (Lake.mod)
  summary (Lake.mod)
```


11. Plot the autocorrelation and partial autocorrelation functions for the residuals from this regression. Interpret them.

```{r}

acf(resid(Lake.mod))
pacf (resid(Lake.mod))

```


12. Fit an autoregressive models to the residuals. Compare the results with your interpretation of the PACF plot.

How to interpretate the PACF. PACF is how my model should look like.

First I separete the residuals of my model into a separate object and then I run the ar(Lake.res)

```{r}

Lake.res <- resid(Lake.mod)

ar(Lake.res) 



```


13. Fit a gls model using a corAR1 correlation structure. Test if the correlation structure is necessary. Is the trend significant?

14.Fit a gls model using a corARMA correlation structure with two AR terms. Is this model an improvement?