---
title: "Mixed Effect Models"
author: "Richard J. Telford"
date: "June 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library (tidyverse)
library (nlme)
```

## R Markdown
Exercises 

1) Import data LME.csv   
id gives the group id, y is the response, ab and AB are predictors
Data needs to be factors, or else the analysis will not treat them correctly. 

```{r}
data.LME <- read.csv("~/Project A1 BIO302/LME.csv")
data.LME <- data.LME %>% mutate(AB = factor (AB), ab = factor(ab)) 

```


2) What is a random effect?

Effects that is particular to an observation (residuals) or group of observations

3) What is a fixed effect?

Effect that is shared by all observations


4) Fit an appropriate a model with y and AB

```{r}
fit.lme <-lme(y~AB,random=~+1|id, data = data.LME)
anova(fit.lme,type="m")
summary (fit.lme)
```


5) Interpret the results

The predictor (outer variable) AB is not significant for the response y. So it doesnt predict the response y. 

6) Fit an appropriate a model with y and ab

```{r}

fit.lme.ab <-lme(y~ab,random=~+1|id, data = data.LME)
anova(fit.lme.ab,type="m")
summary (fit.lme.ab)
```

When you have categorical variables the Intercept is the mean of the first group, and the Value under is th


7) Interpret the results

The predictors ab is signficant and therefore predicts y. 

8) Include both AB and ab as predictors 

```{r}
fit.lme.aAbB <-lme(y~ab+AB,random=~+1|id, data = data.LME)
anova(fit.lme.aAbB,type="m")
summary (fit.lme.aAbB)

data.LME %>%ggplot (aes(x=ab+AB, y = y, )) + geom_boxplot()

```
Intercept Value is the mean for the first group of ab and AB. (1ab, 1AB) = 16.39 as mean. 



9) What is the expected mean for the different levels of AB and ab?



10) Use a linear model with AB and compare the results to question 4



11) Use a linear model with ab and compare the results to question 6

12) Import data length.csv   
id is the individual id, day is the predictor, length is the response

13) Find the best model to explain the data

14) Use a simulation to test the performance of mixed effect models.  

Simulate 10 groups of observations, 5 for each treatment AB, each with 10 individuals. There should be a group error from a normal distribution and an individual error from a normal distribution, and the effect of the predictor. Chose an effect size, individual error and group error that you might expect for your data.

Test how changing the number of individuals within in each group and the number of groups (keeping the total number of observations constant) changes the power of the analysis (the probability of detecting the response).

15) Repeat for an inner variable ab which varies within each group.

16) Consider the implications for experimental design.



