---
title: "GLM & GAM"
author: "Richard J. Telford"
date: "May 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (tidyverse)
library (lmtest)
library (tidyverse)
library (broom)
library (nlme)
library (broom.mixed)
library (mgcv)
```

#Exercises GLM & GAM

##Poisson GLM
1. Import sorich.csv

```{r}
sorich <- read.csv("~/Project A1 BIO302/sorich.csv")


```

2. Fit a linear model for species richness against cover assuming a Normal distribution

3. Study the diagnostic plots

## lm doesnt assume anything else than normal

```{r}

sorich.lm <- lm (nsp ~ cover, data = sorich)
anova (sorich.lm)
summary (sorich.lm)
plot (sorich.lm)
```

4. What distribution is the response?
 
Number of species is count data, so distribution is Poisson.

5. Do an appropriate analysis

```{r}
fit.sorich.glm <- glm(nsp ~ cover, data = sorich, family = poisson)
anova (fit.sorich.glm)
summary (fit.sorich.glm)
plot (fit.sorich.glm)

ggplot(cbind(sorich, fit = fitted(fit.sorich.glm)), aes(x = cover, y = nsp)) +
  geom_point() +
  geom_line(aes(y = fit), colour = "red")



```

6. Check for over-dispersion

Anova gives that Residual Deviance is higher than the DF, which means that variance is to high, so there is a parameter that was not identified. 

7. Interpret the results

There more cover you have, the more number of species. If we fit a quasi model it will give a scale. 

8. How does the width of the confidence interval at cover = 10 change when over-dispersion is allowed for

```{r}
(pred <- predict(fit.sorich.glm, newdata = data.frame(cover = 10), se.fit = TRUE))


#estimate

exp(pred$fit)

#upper 95%ci
exp(pred$fit + 1.96 * pred$se.fit)

#lower 95%ci
exp(pred$fit - 1.96 * pred$se.fit)

#At cover 10 we have 2.19 and that is outside the confidence interval of 95%, which is from 7 to 10. 

```

Cover and grasherb as predictors. Number of species as response. For adding predictors, use +. 
9. Do the grasses differ from the herb, i.e. include the factor grasherb and test its significance?

By adding another predictor, the Resid.Dev is reduced (and as close to DF), so the overdispersion is reduced. But it still not enough.

```{r}

fit.herb.glm <- glm(nsp ~ grasherb + cover, data = sorich, family = poisson)
anova (fit.herb.glm)
summary (fit.herb.glm)
plot (fit.herb.glm)

ggplot(cbind(sorich, fit = fitted(fit.herb.glm)), aes(x = grasherb+ cover, y = nsp)) +
  geom_point() +
  geom_line(aes(y = grasherb+cover), colour = "red")

anova (fit.sorich.glm)
anova(fit.herb.glm)

```



## How much does over-dispersion affect results
1. Use `sample()` to randomise the response in sorich to remove the relationship between the preictor and response.
Sample reorders the vector. We took the column nsp and randomized. We use the dollar sign to indicate that is a colum inside the sorich. We randomized the nsp so we null the relationship between predictor and response.  

```{r}

random.sorich <- tibble (nsp = sample(sorich$nsp), cover = sorich$cover)
random.sorich
```


2. Test if a Poisson GLM with cover as a predictor is significant.

Since the response is random, it might or not be significant. 

```{r}
glm.random.sorich <- glm(nsp ~ cover, data = random.sorich, family = poisson)
anova(glm.random.sorich)
summary (glm.random.sorich)

```


3. Repeat 1000 times and find the distribution of p-values.
Although the data is randomized, I found a significant relation and that is a proof of over dispersion.
The function tidy also gives the intercept value, so I need to filter it out with filter(term =="cover"). With glance, on lm, I don´t need to do that.

```{r}


rerun(.n=1000, tibble (nsp = sample(sorich$nsp), cover = sorich$cover)) %>% 
  map (~ glm(nsp~cover, data = ., family = poisson())) %>% 
  map_df (tidy) %>%  filter(term == "cover") %>%
  ggplot (aes(x=p.value)) + geom_histogram() 


```


## Binomial GAM

1. Open library mgcv
2. Import data pot.csv

```{r}
pot <- read.csv("~/Project A1 BIO302/pot.csv")
pot
```

3. What type of distribution is the response variable?

Binomial 

4. What type of link-function do we use?

Log function 

5. Do an appropriate GLM analysis?
The + I(alt^2) is quadratic, so it will make a curve. 

```{r}

pot.glm <- glm(potalp ~ alt + I(alt^2), data = pot, family = binomial)
anova(pot.glm)
summary (pot.glm)
```


6. Interpret the results?

There is no over dispersion, observing residuals and DF.

7. Do a GAM analysis

```{r}
pot.gam <- gam(potalp ~ s(alt), data = pot, family = binomial)
anova (pot.gam)
```


8. Compare the GLM and GAM models.

```{r}
anova (pot.glm, pot.gam, test ="Chisq")
summary (pot.glm)
summary (pot.gam)

ggplot(cbind(pot, fit.gam = fitted(pot.gam), fit.glm = fitted(pot.glm)), aes(alt, potalp)) +
  geom_point() +
  geom_line(aes(y = fit.gam, colour = "GAM")) +
  geom_line(aes(y = fit.glm, colour = "GLM")) +
  labs(x = "Elevation", y = "P. alpina", colour = "Model")

```

9. Which model would you prefer, and why?

This was the ANOVA output: 

Model 1: potalp ~ alt + I(alt^2)
Model 2: potalp ~ s(alt)
  Resid. Df Resid. Dev     Df Deviance Pr(>Chi)
1    568.00     529.28                         
2    565.94     525.35 2.0608   3.9306   0.1471

the Pr (>Chi) shows that the models are not significantly different, so the choice should go to GLM. Although GAM has a easier plot to read, the GLM is easier to interpret the results and the coeficient, and to measure the highest point in the curve.